---
title: "Peri-INTACT_FISHseq Psuedobulk"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}

here()
```


# Introduction

This analysis aims to determine the cell-specific expression of extracellular matrix (ECM) proteins within the kidney.

# Notes on Analysis (1/10/25)

Within the Kidney Stroma, the following ECM genes exhibit differential expression: **Csgalnact** (fibroblasts), **Dcn** (perivascular fibroblasts), **Ust** (fibroblasts), **Vcan** (inner medulla fibroblasts), **Chsy3** (fibroblasts), and **Cspg4** (contractile cells).

Within the Whole Kidney, the following ECM genes exhibit differential expression: **Chst11** (all but the distal nephron), **Csgalnact1** (endothelium and proximal tubule), **Chst7** (late proximal tubule), **Ust** (fibroblasts and podocytes), **Bgn** (fibroblasts and endothelial cells), **Dse** (contractile cells), and **Chsy3** (fibroblasts and endothelial cells).

# Load Seurat Objects

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

GFP <- readRDS(here("Outputs", "GFP.clean.060424.rds"))

DefaultAssay(GFP) <- "RNA"

DimPlot(GFP)

head(GFP@meta.data)

```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- FindAllMarkers(GFP, only.pos = TRUE, min.pct = 0.25, logfc.threshold = .5)

df

df_unique <- df %>%
  distinct(gene, .keep_all = TRUE)

df_filtered <- df_unique %>%
  filter(!str_starts(gene, "Gm") & !str_ends(gene, "Rik"))

df_top10 <- df_filtered  %>%
  group_by(cluster) %>%
  slice_head(n = 10) %>%
  ungroup()

FeaturePlot(GFP, "Wt1", order = T)

```


# Expression in Mouse Interstial Dataset

```{r echo=TRUE, warning=FALSE, error=FALSE, message=TRUE, fig.align = 'center', fig.width=10}

list <- read.xlsx(here("SpatialGenomics_List_v2.xlsx"), sheet = 1)

genelist <- list$Gene

genelist <- head(genelist, 2)

for (i in genelist) {
  if (!(i %in% rownames(GFP))) {
    message(paste(i, "not found – skipping"))
     }
  
  
  p1 <- VlnPlot(GFP, features = i, group.by = "Special", pt.size = 0.1) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5)) 

  p2 <- FeaturePlot(GFP, features = i, order = TRUE) 

  p3 <- wrap_plots(p1, p2)
  
  print(p3)
}



```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

PB <- AverageExpression(GFP, 
                           group.by = c("Special"), 
                           assays = 'RNA', 
                           slot = "counts",
                           return.seurat = FALSE)

PB <- PB$RNA
PB <- as.matrix(PB)  # Convert to matrix
PB <- as.data.frame(PB) # Convert from matrix to dataframe 


PB2 <- PB %>%
  rownames_to_column(var = "Gene") %>%  # Convert row names to a column
  rowwise() %>%
  mutate(Max_Value = max(c_across(-Gene))) %>%  # Exclude the "Gene" column
  ungroup() %>% 
  mutate(Percent_Rank = percent_rank(Max_Value))


PB3 <- PB2 %>%
 select(Gene, Percent_Rank) 

df3 <- inner_join(df, PB3, by = c("gene" = "Gene"))




```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


DefaultAssay(GFP) <- "SCT"
df2 <- FindAllMarkers(GFP, only.pos = TRUE, min.pct = 0.25, logfc.threshold = .5)

# Convert row names to a column
df4 <- df %>% rownames_to_column(var = "Rowname")
df5 <- df2 %>% rownames_to_column(var = "Rowname")

# Perform inner join by "Rowname"
df6 <- inner_join(df4, df5, by = "Rowname")

# Optionally, remove "Rowname" column and restore row names
df7<- df6 %>% column_to_rownames(var = "Rowname")

ggplot(df7, aes(x = avg_log2FC.x, y = avg_log2FC.y)) +
  geom_point() +  # Adds dots
  labs(x = "Average log2FC (RNA)", y = "Average log2FC (SCT)") +
  theme_minimal() +  # Use minimal theme
  ggtitle("Dotplot of RNA vs SCT") + # Title
  theme_classic()

df8 <- df7 %>% rownames_to_column(var = "Gene")

# Calculate the absolute difference between avg_log2FC.x and avg_log2FC.y
top_genes <- df8 %>%
  mutate(Diff = abs(avg_log2FC.x - avg_log2FC.y)) %>%  # Calculate the difference
  select(Gene, avg_log2FC.x, avg_log2FC.y, Diff) %>%  # Select relevant columns
  arrange(desc(Diff)) %>%  # Sort by largest difference
  head(50)  # Get the top genes with the largest differences

# Print the result
print(top_genes)


```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

